\documentclass[12pt]{article} 
\input{../../custom}
\graphicspath{{figures/}}
\def\showcommentary{1}


\title{In-class exercise for Lecture 2}
\author{}
\date{}


\begin{document}
\maketitle

\subsection*{Instructions}
\begin{itemize}
\item \textbf{Don't look at the solution yet!} This is for your benefit.
\item This exercise must be submitted within 48 hours of the lecture in which it was given. 
\item As long as you do the exercise on time, you get full credit---your performance does not matter.
\item Without looking at the solution, take 5 minutes to try to solve the exercise.
\item Pre-assessment: Write down how correct you think your answer is, from 0 to 100\%.
\item Post-assessment: Now, study the solution and give yourself a ``grade'' from 0 to 100\%.
\item Submit your work on the course website, including the pre- and post- assessments.
\end{itemize}

\subsection*{Exercise}
We write $X\sim\Poisson(\theta)$ if $X$ has the Poisson distribution with rate $\theta>0$, that is, its p.m.f.\ is
$$ p(x|\theta) = \Poisson(x|\theta)= e^{-\theta} \theta^x / x!$$
for $x\in\{0,1,2,\dotsc\}$ (and is $0$ otherwise). Suppose $X_1,\dotsc,X_n\iid\Poisson(\theta)$ given $\theta$, and your prior is
$$ p(\theta) =\Ga(\theta|a,b) =\frac{b^a}{\Gamma(a)} \theta^{a-1} e^{-b\theta}\I(\theta>0) .$$
What is the posterior distribution on $\theta$?

\newpage
\vfill
\rotatebox{180}{
\begin{minipage}[t][\textheight][t]{\textwidth}
\subsection*{Solution}
\scriptsize
Since the data is independent given $\theta$, the likelihood factors and we get
\begin{align*}
p(x_{1:n}|\theta) & = \prod_{i = 1}^n p(x_i|\theta) \\
& = \prod_{i = 1}^n e^{-\theta} \theta^{x_i} / x_i! \\
& \underset{\theta}{\propto} e^{-n\theta} \theta^{\sum x_i}.
\end{align*}
Thus, using Bayes' theorem,
\begin{align*}
p(\theta|x_{1:n}) &\propto p(x_{1:n}|\theta) p(\theta) \\
& \propto e^{-n\theta} \theta^{\sum x_i} \theta^{a-1} e^{-b\theta}\I(\theta>0) \\
& \propto e^{-(b+n)\theta} \theta^{a+\sum x_i-1}\I(\theta>0) \\
& \propto \Ga\big(\theta\mid a+\textstyle\sum x_i,\,b+n\big).
\end{align*}
Therefore, since the posterior density must integrate to $1$, we have
$$p(\theta|x_{1:n}) =\Ga\big(\theta\mid a+\textstyle\sum x_i,\,b+n\big).$$
\end{minipage}}

\end{document}






